---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# states

<!-- badges: start -->
[![CRAN versions](http://www.r-pkg.org/badges/version/states)](https://cran.r-project.org/package=states)
[![R-CMD-check](https://github.com/andybega/states/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/andybega/states/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/andybega/states/graph/badge.svg)](https://app.codecov.io/gh/andybega/states)
<!-- badges: end -->

Create country-year/month/day panels consistent with the COW or Gleditsch & Ward lists of independent states. I mainly use this for merging different data sources: 

1. Create a master template that reflects one of the independent states lists. 
2. For each data source, normalize to a copy of the master template. Doing this by source makes it easier to identify and address issues like missing values or observation for non-independent states. 
3. In the end, merge everything together. Since all the inputs are already normalized to a proper state panel list, there should be no issues.

What the package does:

- It contains the Gleditsch and Ward (G&W) as well as the Correlates of War (COW) state system membership lists. 
```{r}
library("states")
    
data(gwstates)
data(cowstates)
```

- Search them with `sfind`, this can be helpful for manual coding:

```{r}
sfind(260)[, 1:6]
sfind("German")[, 1:6]
```

- You can use it to build a country-year template that matches either the COW or G&W state lists. 
```{r}
countries <- state_panel(1991, 2001)
    
str(countries)
```

- Visualize missing/non-independent cases in your data. See this related [blog post for more details](https://www.andybeger.com/blog/2016-09-13-data-management/).

```{r, fig.height = 5}
data("polity")
str(polity)
    
plot_missing(polity, x = "polity", ccode = "ccode", statelist = "COW")
```

For the underlying data (missing info requires date input):

```{r}
polity$date <- as.Date(paste0(polity$year, "-01-01"))
mm <- missing_info(polity, x = "polity", ccode = "ccode", 
                   time = "date", period = "year",
                   statelist = "COW")
head(mm)
```

## Install

The package is on CRAN and can be installed with:

```r
install.packages("states")
```

Or to install from GitHub:

```r
library("remotes")
remotes::install_github("andybega/states")
```

## Citations

For the [Gleditsch and Ward (G&W) state data](http://ksgleditsch.com/statelist.html):

Gleditsch, Kristian S. & Michael D. Ward. 1999. "Interstate System Membership: A Revised List of the Independent States since 1816." International Interactions 25: 393-413.

For the [Correlates of War (COW) state data](https://correlatesofwar.org/data-sets/state-system-membership/):

Correlates of War Project. 2017. "State System Membership List, v2016." Online, https://correlatesofwar.org
