---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# states

[![CRAN versions](http://www.r-pkg.org/badges/version/states)](https://cran.r-project.org/package=states)
[![Travis-CI Build Status](https://travis-ci.org/andybega/states.svg?branch=master)](https://travis-ci.org/andybega/states)

Create country-year/month/day panels consistent with the COW or Gleditsch & Ward lists of independent states. I mainly use this for merging different data sources: 

1. Create a master template that reflects one of the independent states lists. 
2. For each data source, normalize to a copy of the master template. Doing this by source makes it easier to identify and address issues like missing values or observation for non-independent states. 
3. In the end, merge everything together. Since all the inputs are already normalized to a proper state panel list, there should be no issues.

What the package does:

1. It contains the Gleditsch and Ward (G&W) as well as the Correlates of War (COW) state system membership lists. 
    ```{r}
    library("states")
    
    data(gwstates)
    data(cowstates)
    ```
    Search them with `sfind`, this can be helpful for manual coding:
    ```{r}
    sfind(260)[, 1:6]
    sfind("German")[, 1:6]
    ```
2. You can use it to build a country-year template that matches either the COW or G&W state lists. 
    ```{r}
    countries <- state_panel("1991-01-01", "2010-01-01", by = "year")
    
    str(countries)
    ```
3. Visualize missing/non-independent cases in your data. See this related [blog post for more details](https://andybeger.com/2016/09/14/data-management/).
    ```{r, fig.height = 5}
    data("polity")
    str(polity)
    polity$date <- as.Date(paste0(polity$year, "-12-31"))
    
    plot_missing("polity", polity, "ccode", "date", "year", statelist = "COW")
    ```

## Install

The package is on CRAN and can be installed with:

```r
install.packages("states")
```

Or to install from GitHub:

```r
library(devtools)
install_github("andybega/states")
```

## Citations

For the [Gleditsch and Ward (G&W) state data](http://privatewww.essex.ac.uk/~ksg/statelist.html):

Gleditsch, Kristian S. & Michael D. Ward. 1999. "Interstate System Membership: A Revised List of the Independent States since 1816." International Interactions 25: 393-413.

For the [Correlates of War (COW) state data](http://www.correlatesofwar.org/data-sets/state-system-membership):

Correlates of War Project. 2017. "State System Membership List, v2016." Online, http://correlatesofwar.org

