---
title: "Country code differences to look out for"
author: "Andreas Beger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("states")
library("magrittr")
library("dplyr")
```

Notes on common country code differences when comparing and merging different data sources. 

## Between COW and G&W lists

```{r}

data("cowstates")
data("gwstates")

```

How many COW country names are in G&W country names?

```{r}
table(cowstates$country_name %in% gwstates$country_name)
```

How many G&W country names are in COW country names?

```{r}
table(gwstates$country_name %in% cowstates$country)
```

There is more apparent overlap when we focus on country codes:

```{r}
cowstates %>% filter(!cowcode %in% gwstates$gwcode) %>%
  knitr::kable(caption = "Country codes in COW but not in G&W")

gwstates %>% filter(!gwcode %in% cowstates$cowcode) %>%
  knitr::kable(caption = "Country codes in G&W but not in COW")
```


This is deceiving however. There are several country codes that refer to different countries in COW and G&W. Also, which code is retained for states after a split or merge is different in several cases. 

- Germany after 1991 unification: COW uses 255, G&W continues 260
- Yemen: COW has 678 for Yemen from 1926 to 1990, then 679 after unification; G&W uses 678 throughout, but independence is earlier. 
- Kiribati: 970 in G&W, 946 in COW
- Nauru: 971 in G&W, 970 in COW
- Tuvalu: 973 in G&W, 947 in COW
- Tonga: 972 in G&W, 955 in COW


## G&W and WDI

Some example code for converting WDI data to G&W:

```r
WDI(...) %>%
  mutate(gwcode = countrycode(iso2c, "iso2c", "cown")) %>%
  mutate(gwcode = case_when(
           iso2c=="RS" ~ 340L,
           iso2c=="XK" ~ 347L,
           iso2c=="SS" ~ 626L,
           gwcode==255L & year >= 1949 ~ 260L,
           gwcode==679L ~ 678L,
           TRUE ~ gwcode
         ))
```
